Delayed dichromatism as a convenient tool to disentangle the effects of survival and productivity on the population dynamics in waterfowl

### 2024-05-29 Code/Draft_ruddy_duck_vital_rates.Rmd
keywords modif

# Materials & methods {#MM}

@@

## Validating the "apparent sex ratio" method {#VAL}

@@

We used the Bayesian framework to implement all submodels presented in Section \@ref(MM). The Bayesian framework is both straightforward and efficient in propagating error through the parameters. We used uninformative priors for all parameters. Since the maximum growth rate is a life history trait expected to be stable between populations of the same species, we used an uninformative hierarchical prior for this parameter. We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, and discarded the first 2,000 samples as burn-in. We checked the chain convergence using the Gelman and Rubin convergence diagnostic (R\<1.1, @Gelman1992). The models were fitted using NIMBLE [@Valpine2017] run from R [@RCT2022]. The values **X[Y; Z]** reported in Section \@ref(RES) are the medians and the corresponding limits of the 95% confidence interval of the posterior distributions. We preferred the median to the mean because it is more robust to skewed distributions.

# Results {#RES}

@@

Recruitment rates ranged between 0.09 [0.04; 0.13] and 0.75 [0.63; 0.87]. All estimates were above 0, which is not inconsistent with reality. Furthermore, the maximum recruitment rate was 0.68 [0.36; 0.78] (see Section \@ref(MAX)), and no estimate was significantly outside the range of recruitment rates defined without immigration [0; 0.78]. No trend was observed for the GB population, but the recruitment rate decreased for the FR population, although it was more noisy than the proportion of immatures. As with the survival rate, the GB population also showed lower recruitment rate values with less variability than the FR population.

@@

## Assessing the demographic response to culling strategies {#MAX}

@@

Despite the low number of sex- and age-determined culled individuals, the proxies of culling pressure showed significant changes over time (Figure \@ref(fig:sup)). Therefore, we categorised culling pressure according to this intensity. High culling pressure occurred from 2006 onwards in both populations (although the signal was noisy). However, despite the similar level of culling pressure, the culling strategies of the two populations differed considerably, see Section \@ref(HARV). Under high culling pressure, the growth rate decreased to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for GB and FR, corresponding to a decrease of 44% per year and 3% per year, respectively. This result shows that the GB population declined significantly during the high culling pressure, while the FR population was just stabilised.

>>>>

# Materials & methods

@@

## Validating the "apparent sex ratio" method

@@

We used the Bayesian framework to implement all submodels presented in Section \@ref(MM). The Bayesian framework is both straightforward and efficient in propagating error through the parameters. We used uninformative priors for all parameters. Since the maximum growth rate is a life history trait expected to be stable between populations of the same species, we used an uninformative hierarchical prior for this parameter. We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, and discarded the first 2,000 samples as burn-in. We checked the chain convergence using the Gelman and Rubin convergence diagnostic (R\<1.1, @Gelman1992). The models were fitted using NIMBLE [@Valpine2017] run from R [@RCT2022]. The values **X[Y; Z]** reported in Results section are the medians and the corresponding limits of the 95% confidence interval of the posterior distributions. We preferred the median to the mean because it is more robust to skewed distributions.

# Results

@@

Recruitment rates ranged between 0.09 [0.04; 0.13] and 0.75 [0.63; 0.87]. All estimates were above 0, which is not inconsistent with reality. Furthermore, the maximum recruitment rate was 0.68 [0.36; 0.78] (see 3rd Results section), and no estimate was significantly outside the range of recruitment rates defined without immigration [0; 0.78]. No trend was observed for the GB population, but the recruitment rate decreased for the FR population, although it was more noisy than the proportion of immatures. As with the survival rate, the GB population also showed lower recruitment rate values with less variability than the FR population.

@@

## Assessing the demographic response to culling strategies

@@

Despite the low number of sex- and age-determined culled individuals, the proxies of culling pressure showed significant changes over time (Figure \@ref(fig:sup)). Therefore, we categorised culling pressure according to this intensity. High culling pressure occurred from 2006 onwards in both populations (although the signal was noisy). However, despite the similar level of culling pressure, the culling strategies of the two populations differed considerably, see 3rd Materials & Methods section. Under high culling pressure, the growth rate decreased to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for GB and FR, corresponding to a decrease of 44% per year and 3% per year, respectively. This result shows that the GB population declined significantly during the high culling pressure, while the FR population was just stabilised.
